\Huge Conferences in an Era of Expensive Carbon

### 2019-12-21 main.tex
A few nites

\newcommand{\bcp}[1]{\ifdraft{\bf [bcp: #1]}\fi}
\newcommand{\mwh}[1]{\ifdraft{\bf [mwh: #1]}\fi}
\newcommand{\COtwoE}{CO$_2$e}

@@

\title{\Huge Conferences in the Era of Expensive Carbon}
% \title{\Huge ACM Conferences and Cost of Carbon}
\author{Benjamin C. Pierce (University of Pennsylvania) \\
Michael Hicks (University of Maryland) \\
Crista Lopes (University California, Irvine) \\
Jens Palsberg (University of California, Los Angeles)}

@@

Indeed, scientific societies arguably have an especially important role to
play, since, for many of their members, travel to conferences may be a
substantial or even dominant part of their individual contribution to
climate change.  A single round-trip flight from Philadelphia to Paris emits
the equivalent of about 1.7 tons of carbon dioxide (\COtwoE) per
passenger\iflater\bcp{Citation?}\fi, a significant fraction of the
{\em total} yearly emissions for an average resident of the US (16.5 tons)
or Europe (7 tons)~\cite{emissions}.  
Moreover, these emissions have no near-term technological
fix, since jet fuel is difficult to replace with renewable energy
sources~\cite{elec-air}.
% \iflater\bcp{Trimmed this sentence: Reducing the \COtwoE{} footprint
% of conferences will require a reduction in the footprint of the attendees'
% air travel.}\fi

How should ACM respond to these facts?  Two years ago, \mwh{Three, now?} ACM's Special
Interest Group on Programming Languages (SIGPLAN) convened an ad hoc Climate
Committee to research this question~\cite{ClimateCommittee}.  After
investigating many options~\cite{ClimateCommitteReport} and intensive and
wide-ranging discussions, we are putting forward two concrete
proposals. First, all ACM 
conferences should publicly {\em account} for the {\COtwoE} emitted as a
result of putting them on---in particular, from travel to
the conference. Second, ACM should put a \emph{price} on 
carbon in conference budgets, to create pressure on organizers to reduce
their footprints.

@@

Many organizations, including companies such as Google, Dell, Microsoft,
General Motors, Delta Airlines, Lyft, and Expedia, as well as universities,
academic societies, and even energy companies such as Exxon, now use
carbon offsets in order to reduce their 
net footprint. ACM conferences can do the same, and the purchases can be
mentioned in the public accounting we are proposing.  In fact, ACM recently
decided that conferences could offer attendees the option to pay for an
offset during registration.  Other scientific societies are going further,
rolling carbon offsets into the registration fees for all
participants.\iflater\bcp{In a later draft, we may be able to say that ICML,
NIPS, etc. are a case in point.  Ask Alexandre Lacoste about this.}\fi

@@

If you've already thought about all of these, then that's great.  If not, please do take them as a set of potential
challenges to consider (all voluntary... as I intend them only as suggestions).
\end{quote}
\item 
About Crista's figure:
BCP: Would be great to supplement with a more refined analysis of registrations for each of these conferences; then we could add a "center of mass" dot for each of the conferences to the map. This would still not tell the whole story, since it's possible (likely) that the registration totals would have been different if the conferences had been located differently. But it would help tell the story.
\item 
Also about Crista's figure:
\begin{itemize}
\item 
BCP: We should quantify what the red and blue dots mean -- i.e., give readers an idea what overall difference a change of location would make to carbon footprint.
\item Mike: The dot colors mean per-capita footprint. Could we make their size represent the total footprint? Far-flung places might be high per-capita, but lower overall.
\item Crista: The order by total footprint is very similar to the normalized one. Here are the 5 top offenders by that criteria: ICFP-Japan, POPL-LA, POPL-SD, SPLASH-Vancouver, PLDI-Barcelona. There are some changes in the order of the conferences in the middle. I hesitated which plot to draw (each has positives and negatives) but decided this one would be less problematic. I don't think we have space for two, and overlaying the 2 data series will be confusing...
\item BTW, there's a LOT more that can be said about this data... But I think that will distract from the main message of this article. We need to chose whether this 2-page article is about the data analysis of our conferences or about a call for action. Maybe we can write a data analysis paper, upload it on arXiv, and link to it from this article?
\item
For example: POPL-Paris was the absolute winner by all criteria: it had the highest number of participants, the second-lowest normalized carbon footprint, and was mid-point in total carbon footprint, even with all those participants.
This is just one tidbit; there are many more interesting facts in this data.
\end{itemize}
\item 
From Michael Coblenz:
What about research on energy-efficient computing? The impact of this research might be far more significant than that of mitigating our direct carbon emissions, and this is more in line with the core skills of the community.
Also, what about research on telepresence and virtual meetings? There's obviously a lot of work in this area, but we may need more in order to (a) let us effectively trade off options, e.g. virtual PC meeting vs in-person; (b) improve the state of the art to make remote meetings more effective.
Consider also discussing the diversity implications on travel. Reducing travel expectations might make the field more attractive to women, who may face higher social expectations to stay home (and are more likely to be single parents). Surely there is a good article to cite here. I found https://journals.sagepub.com/doi/abs/10.1177/0950017006066999 but maybe there is something better. I'm looking…
\item 
Consider citing the new IEA report as another piece of evidence that action is needed
\item 
From Sophia: Does the article need an abstract to "set the reader's expectations"? I expected that there would be a list of actions in the end, or recommended resolutions. Instead, the article is more about what some communities do, and where individuals may be involved.  Perhaps say: "In this article we give a short overview of how different SCM communities tackle this problem, what thought are relevant to it, and conclude with suggestions on how to get involved". It breaks the flow, but it clarifies what the article is about.
\item 
Is there a ``more authoritative'' source for price of carbon offsets than
our carbon report?
\item 
Check where we've used ``carbon'' (or CO2) and where ``greenhouse gases''.  Gloss CO2 or CO2e as appropriate.
\end{itemize}
\fi

>>>>

\usepackage{xcolor}
\definecolor{darkred}{rgb}{0.7, 0.0, 0.0}
\newcommand{\bcp}[1]{\ifdraft{\bf\textcolor{darkred}{[\bf bcp: #1]}}\fi}
\newcommand{\mwh}[1]{\ifdraft{\bf\textcolor{darkred}{[\bf mwh: #1]}}\fi}
\newcommand{\COtwoE}{CO$_2$e}

@@

\title{\Huge Conferences in an Era of Expensive Carbon}
% \title{\Huge ACM Conferences and Cost of Carbon}
\author{Benjamin C. Pierce (University of Pennsylvania) \\
Michael Hicks (University of Maryland) \\
Crista Lopes (University California, Irvine) \\
Jens Palsberg (University of California, Los Angeles)}

@@

Indeed, scientific societies arguably have an especially important role to
play, since, for many of their members, travel to conferences may be a
substantial or even dominant part of their individual contribution to
climate change.  A single round-trip flight from Philadelphia to Paris emits
the equivalent of about 1.7 tons of carbon dioxide (\COtwoE) per
passenger\iflater\bcp{Citation?  Yes, to the same carbon calculator that we
  recommend elsewhere!}\fi, a significant fraction of the
{\em total} yearly emissions for an average resident of the US (16.5 tons)
or Europe (7 tons)~\cite{emissions}.
Moreover, these emissions have no near-term technological
fix, since jet fuel is difficult to replace with renewable energy
sources~\cite{elec-air}.
% \iflater\bcp{Trimmed this sentence: Reducing the \COtwoE{} footprint
% of conferences will require a reduction in the footprint of the attendees'
% air travel.}\fi

How should ACM respond to these facts?  In 2016, ACM's Special
Interest Group on Programming Languages (SIGPLAN) convened an ad hoc Climate
Committee to research this question~\cite{ClimateCommittee}.  After
investigating many options~\cite{ClimateCommitteReport} and intensive and
wide-ranging discussions, we are putting forward two concrete
proposals. First, all ACM
conferences should publicly {\em account} for the {\COtwoE} emitted as a
result of putting them on---in particular, from travel to
the conference. Second, ACM should put a \emph{price} on
carbon in conference budgets, to create pressure on organizers to reduce
their footprints.

@@

Many organizations, including companies such as Google, Dell, Microsoft,
General Motors, Delta Airlines, Lyft, and Expedia, as well as universities,
academic societies, and even energy companies such as Exxon, now use
carbon offsets in order to reduce their
net footprint. ACM conferences can do the same, and the purchases can be
mentioned in the public accounting we are proposing.  In fact, ACM recently
decided that conferences could offer attendees the option to pay for an
offset during registration.  Other scientific societies are going further,
rolling a carbon offset into the registration fee for every participant.

@@

If you've already thought about all of these, then that's great.  If not, please do take them as a set of potential
challenges to consider (all voluntary... as I intend them only as suggestions).
\end{quote}
\item
About Crista's figure:
BCP: Would be great to supplement with a more refined analysis of registrations for each of these conferences; then we could add a "center of mass" dot for each of the conferences to the map. This would still not tell the whole story, since it's possible (likely) that the registration totals would have been different if the conferences had been located differently. But it would help tell the story.
\item
Also about Crista's figure:
\begin{itemize}
\item
BCP: We should quantify what the red and blue dots mean -- i.e., give readers an idea what overall difference a change of location would make to carbon footprint.
\item Mike: The dot colors mean per-capita footprint. Could we make their size represent the total footprint? Far-flung places might be high per-capita, but lower overall.
\item Crista: The order by total footprint is very similar to the normalized one. Here are the 5 top offenders by that criteria: ICFP-Japan, POPL-LA, POPL-SD, SPLASH-Vancouver, PLDI-Barcelona. There are some changes in the order of the conferences in the middle. I hesitated which plot to draw (each has positives and negatives) but decided this one would be less problematic. I don't think we have space for two, and overlaying the 2 data series will be confusing...
\item BTW, there's a LOT more that can be said about this data... But I think that will distract from the main message of this article. We need to chose whether this 2-page article is about the data analysis of our conferences or about a call for action. Maybe we can write a data analysis paper, upload it on arXiv, and link to it from this article?
\item
For example: POPL-Paris was the absolute winner by all criteria: it had the highest number of participants, the second-lowest normalized carbon footprint, and was mid-point in total carbon footprint, even with all those participants.
This is just one tidbit; there are many more interesting facts in this data.
\end{itemize}
\item
From Michael Coblenz:
What about research on energy-efficient computing? The impact of this research might be far more significant than that of mitigating our direct carbon emissions, and this is more in line with the core skills of the community.
Also, what about research on telepresence and virtual meetings? There's obviously a lot of work in this area, but we may need more in order to (a) let us effectively trade off options, e.g. virtual PC meeting vs in-person; (b) improve the state of the art to make remote meetings more effective.
Consider also discussing the diversity implications on travel. Reducing travel expectations might make the field more attractive to women, who may face higher social expectations to stay home (and are more likely to be single parents). Surely there is a good article to cite here. I found https://journals.sagepub.com/doi/abs/10.1177/0950017006066999 but maybe there is something better. I'm looking…
\item
Consider citing the new IEA report as another piece of evidence that action is needed
\item
From Sophia: Does the article need an abstract to "set the reader's expectations"? I expected that there would be a list of actions in the end, or recommended resolutions. Instead, the article is more about what some communities do, and where individuals may be involved.  Perhaps say: "In this article we give a short overview of how different SCM communities tackle this problem, what thought are relevant to it, and conclude with suggestions on how to get involved". It breaks the flow, but it clarifies what the article is about.
\item
Is there a ``more authoritative'' source for price of carbon offsets than
our carbon report?
\item
Check where we've used ``carbon'' (or CO2) and where ``greenhouse gases''.  Gloss CO2 or CO2e as appropriate.
\end{itemize}
\fi
